SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-10-21 maintex.tex
moved Cheng Li to acknowledgements
minor edits up to Sec 2

When the gas-inflows impact upon the gases in the nucleus of galaxy a burst of new star formation in triggered \citep{Barnes:1996, Mihos:1996}. These gas inflows will also bring metal-poor gases from the disk into the center of the galaxy which can dilute the central metallicity \citep{Rupke:2010, Perez:2011, Scudder:2012}. The gas-inflows may also be able to make it into very center of the galaxy and trigger supermassive black hole (SMBH) accretion \citep{Capelo:2017}.

@@

Indeed, \citet{Barrera-Ballesteros:2015} used the CALIFA (Calar Alto Legacy Integral Field Area) survey to study a sample of 103 paired galaxies by varying the size of the aperture through which the \ewha\ is extracted from. In the study, \citet{Barrera-Ballesteros:2015} found a moderate enhancement to the sSFR in the centers of paired galaxies and a moderate suppression to the sSFR in outskirts of the paired galaxies. 
%
\citet{Pan:2019} used the MaNGA survey to study radial profiles of a sample of 205 paired galaxies. The enhancement to the sSFR was shown to be the strongest in the centers of the paired galaxies. This central enhancement linearly fell with increasing galactocentric radii; however, a moderate enhancement to the sSFR remain in the outskirts of the galaxies. \citet{Pan:2019} further studied the paired galaxies as a function of merger stage; from well separated pairs to post-merger galaxies. Across the different merger stages, the sSFR enhancement was greatest in close pairs with tidal features and in post-merger galaxies. This was in agreement with previous hydrodynamical simulations which showed that a burst of star formation is triggered after the first pericenter and as the two galaxies begin to coalesce \citep{Scudder:2012}. 
%
The radial profile of sSFR in post-merger galaxies has also been studied with the MaNGA survey by \citet{Thorp:2019}. The post-merger galaxies were shown to have a strong enhancement to the sSFR in their centers as well as a moderate enhancement in their outskirts.

Where previous studies on the radial profile of the sSFR offsets in paired galaxies have focused on studying the profiles as a function of interaction stage, we will focus on the radial profile as a function of the stellar mass, projected separation, and mass ratio. As mentioned previously, these parameters have been covered by studies restricted to the nuclear region of the paired galaxies. With the MaNGA survey, we will be able to expand upon these previous studies in greater spatial detail. We will be able to see how these three parameters affect the both the level of the sSFR offsets in the centers of the paired galaxies and the offsets in the outskirts of the galaxies. We will also be able to study if any of three parameters influence the gradient of the sSFR enhancement profiles or if the gradient is preserved between different configurations.

In our previous work, we built a sample of paired galaxies in the MaNGA survey where both components of the pair were contained within field of view of a single integral field unit \citep[][hereafter \citetalias{Fu:2018}]{Fu:2018}. We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we will use a version of this sample updated for the recent MaNGA Data Release 16 (DR16) \citep{Ahumada:2020} and we will further supplement this with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $< z <$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

\begin{equation}\label{eq:blue}
NUV-r = 3.1682 - 0.16 (\mathcal{M}_{\rm i}+18)
\end{equation}
\begin{equation}\label{eq:red}
NUV-r = 4.7866 - 0.04 (\mathcal{M}_{\rm i}+18)
\end{equation}

Where $NUV-r$ is the color from SDSS's k-corrected absolute magnitude and $\mathcal{M}_{\rm i}$ is the i-band magnitude from the NSA catalog. We further apply an \ewha\ $\ge$ 3\AA\ cut on galaxy spectra extracted from a 1.0 \reff\ aperture to ensure that all passive galaxies are removed from the sample \citep{Cid-Fernandes:2011}. We further use the BPT diagnostic \citep{Baldwin:1981} to remove galaxies with an AGN in their centers from our sample. We use the emission line ratios, log([O\textsc{III}]/H$\beta$) and log([N\textsc{II}]/H$\alpha$), extracted from a 1.3 kpc aperture along with the maximum starburst line from \citet{Kewley:2001} as the demarcation between star forming branch and the AGN branch.

@@

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction is parameterized as $R_{\rm V}$ $\equiv$ $A_{\rm V}/E(B-V)$ $=$ 3.1, where we estimate the value of the V-band extinction, A$_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination.

@@

For the outside-IFU sample we use the stellar masses calculated from the r-band elliptical Petrosian apertures from the NSA catalog. Most of the companion galaxies in the inside-IFU sample are not covered by either the NSA catalog or \citet{Simard:2011}. Because of this, the inside-IFU galaxies are left out of this analysis.

@@

% funding
J.S. and H.F. acknowledge support from the National Science Foundation (NSF) grant AST-1614326. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS acknowledges support and resources from the Center for High-Performance Computing at the University of Utah. The SDSS web site is www.sdss.org.

>>>>

When the gas-inflows impact upon the gases in the nucleus of galaxy a burst of new star formation in triggered \citep{Barnes:1996, Mihos:1996}. These gas inflows will also bring metal-poor gases from the disk into the center of the galaxy which can dilute the central metallicity \citep{Rupke:2010, Perez:2011, Scudder:2012}. The gas-inflows may also be able to reach the very center of the galaxy and trigger supermassive black hole (SMBH) accretion \citep{Capelo:2017}.

@@

Indeed, \citet{Barrera-Ballesteros:2015} used the CALIFA (Calar Alto Legacy Integral Field Area) survey to study a sample of 103 paired galaxies by varying the size of the aperture through which the \ewha\ is extracted from. In this study, they found a moderate enhancement to the sSFR in the centers of paired galaxies and a moderate suppression to the sSFR in outskirts of the paired galaxies. 
%
\citet{Pan:2019} used the SDSS-IV MaNGA survey to study radial profiles of a sample of 205 paired galaxies. The enhancement to the sSFR was shown to be the strongest in the centers of the paired galaxies. This central enhancement linearly fell with increasing galactocentric radii; however, a moderate enhancement to the sSFR remain in the outskirts of the galaxies. \citet{Pan:2019} further studied the paired galaxies as a function of merger stage; from well separated pairs to post-merger galaxies. Across the different merger stages, the sSFR enhancement was greatest in close pairs with tidal features and in post-merger galaxies. This was in agreement with previous hydrodynamical simulations which showed that a burst of star formation is triggered after the first pericenter and as the two galaxies begin to coalesce \citep{Scudder:2012}. 
%
The radial profile of sSFR in post-merger galaxies has also been studied with the MaNGA survey by \citet{Thorp:2019}. The post-merger galaxies were shown to have a strong enhancement to the sSFR in their centers as well as a moderate enhancement in their outskirts.

Where previous studies on the radial profile of the sSFR offsets in paired galaxies have focused on studying the profiles as a function of interaction stage, we will focus on the radial profile as a function of the stellar mass, projected separation, and mass ratio. As mentioned previously, these parameters have been covered by studies restricted to the nuclear region of the paired galaxies. With the MaNGA survey, we will be able to expand upon these previous studies in greater spatial detail. We will be able to see how these three parameters affect both the level of the sSFR offsets in the centers of the paired galaxies and the offsets in the outskirts of the galaxies. We will also be able to study if any of three parameters influence the gradient of the sSFR enhancement profiles or if the gradient is preserved between different configurations.

In our previous work using the MaNGA data included in the 14th Public Data Release \citep[DR14;][]{Abolfathi:2018}, we built a sample of close galaxy pairs where both components of the pair were contained within the field of view of a single integral field unit \citep[][hereafter \citetalias{Fu:2018}]{Fu:2018}. We found that approximately 5.7\% of the MaNGA galaxies have a companion galaxy contained within the field-of-view of a single IFU. In this work, we update this sample and supplement it with a sample of companion galaxies identified outside the field-of-view of the MaNGA IFUs.

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies in the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $< z <$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_i}$ $<$ -24.0, where $\mathcal{M_i}$ is the rest frame $i$-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 \reff\ and 2.5 \reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

\begin{equation}\label{eq:blue}
NUV-r = 3.1682 - 0.16 (\mathcal{M}_i+18)
\end{equation}
\begin{equation}\label{eq:red}
NUV-r = 4.7866 - 0.04 (\mathcal{M}_i+18)
\end{equation}

Where $NUV-r$ is the color from SDSS's $k$-corrected absolute magnitude and $\mathcal{M}_i$ is the $i$-band magnitude from the NSA catalog. We further apply an \ewha\ $\ge$ 3\AA\ cut on galaxy spectra extracted from a 1.0~\reff\ aperture to ensure that all passive galaxies are removed from the sample \citep{Cid-Fernandes:2011}. Finally, we use the BPT diagnostic \citep{Baldwin:1981} to remove galaxies with possible AGN in their centers. Specifically, we use the emission line ratios, log([O~\sc{iii}]/H$\beta$) and log([N~\sc{ii}]/H$\alpha$), extracted from a 1.3~kpc aperture along with the maximum starburst line of \citet{Kewley:2001} as the demarcation between the star-forming branch and the AGN branch.

@@

%%% Reddening
We correct the emission lines for reddening using the extinction curve from \citet{Cardelli:1989} with updated coefficients from \citet{ODonnell:1994}. The extinction is parameterized as $R_V$ $\equiv$ $A_V/E(B-V)$ $=$ 3.1, where we estimate the value of the $V$-band extinction, $A_V$, by comparing the H$\alpha$/H$\beta$ ratio to the expected value of 2.85 for case-B recombination.

@@

For the outside-IFU sample we use the stellar masses calculated from the $r$-band elliptical Petrosian apertures from the NSA catalog. Most of the companion galaxies in the inside-IFU sample are not covered by either the NSA catalog or \citet{Simard:2011}. Because of this, the inside-IFU galaxies are left out of this analysis.

@@

% funding
We thank Cheng Li for our useful discussion on how the sSFR enhancement in galaxy pairs relates to their stellar mass and mass ratio. J.S. and H.F. acknowledge support from the National Science Foundation (NSF) grant AST-1614326. Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P. Sloan Foundation, the U.S. Department of Energy Office of Science, and the Participating Institutions. SDSS acknowledges support and resources from the Center for High-Performance Computing at the University of Utah. The SDSS web site is www.sdss.org.
